---
title: "Data Carpentry"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "middle", fig.width = 9, fig.height = 6)
```

```{r}
library(tidyverse)
library(sf)
library(geojsonio)
library(mxmaps)
library(lubridate)
library(knitr)
```

```{r}
prettify <- function(df, rowstart = 1, rowstop = 3, ...){
  df %>% 
    slice(rowstart:rowstop) %>%
    mutate_if(is.Date, as.character) %>%
    mutate_if(is.difftime, as.character) %>%
    mutate_if(is.character, replace_na, "") %>% 
    kable(format = "markdown", ...)
}
```


```{r}
base_geom_sf <- maps::map(fill = TRUE, plot = FALSE) %>% 
  st_as_sf() %>%
  as_tibble() %>%
  st_as_sf() %>% 
  st_crop(xmin = -121, ymin = 13, xmax = -82, ymax = 32)
```


# Radiological Theft Events

The manual cataloguing of radiological theft events was only possible through the persistent efforts of [Francisco "Paco" Parada](https://twitter.com/paco_parada).

```{r}
raw_events_df <- read_csv("data/raw_events.csv")
```

# Flat Data

## Crimes

This incredibly useful Mexican crime data was made possible by [Diego Valle](https://twitter.com/diegovalle). It can be found [here](https://elcri.men/en/datos.html).

```{r}
temp_file <- tempfile(fileext = ".gz")
download.file("https://elcri.men/data/nm-fuero-comun-municipios.csv.gz",
              destfile = temp_file)
```

```{r}
crimes_df <- read_csv(gzfile(temp_file)) %>% 
  mutate(count = as.numeric(count)) %>%
  filter(count != 0) %>%
  distinct() %>% 
  filter(subtipo %in% c("ROBO DE VEHÍCULO AUTOMOTOR",
                        "ROBO EN TRANSPORTE INDIVIDUAL",
                        "ROBO EN TRANSPORTE PÚBLICO COLECTIVO",
                        "ROBO EN TRANSPORTE PÚBLICO INDIVIDUAL",
                        "ROBO A TRANSPORTISTA"  )) %>%
  mutate(INEGI = str_mxmunicipio(state_code, mun_code)) %>%
  mutate(date = date %>%
           paste0( "-01") %>%
           as.Date(format = "%Y-%m-%d")
         ) %>%
  select(date, INEGI, tipo, subtipo, modalidad, count) %>% 
  distinct()
```

```{r}
crimes_df %>% 
  prettify()
```

## Demographics

```{r}
demographics_df <- df_mxmunicipio %>% 
  as_tibble() %>% 
  mutate(INEGI = str_mxmunicipio(state_code, municipio_code)) %>% 
  select(INEGI, municipio_name, pop, pop_male, pop_female,
         metro_area) %>% 
  mutate(is_metro = if_else(is.na(metro_area), FALSE, TRUE)) %>% 
  distinct()
```

```{r}
demographics_df %>% 
  prettify()
```

# Spatial Data

## Radiological Theft Geometries

```{r}
new_names <- list(
  timeUTC = "UTC -6",
  eventID = "EventID",
  location = "Location",
  date = "Dates",
  timeLocal = "localtimerobbery",
  daysToRecovery = "Reco_time",
  theftRecovery = "Event",
  reportNumber = "reportnum",
  licensee = "Licensee",
  lat = "Lat",
  long = "Long",
  distanceTheftRecovery = "distance",
  category = "Category",
  typeIEC = "Type_IEC",
  thesisGroup = "Thesis_grouping",
  source = "Source",
  isotope = "Isotope",
  sourceActivityCi = "Source Activity Ci",
  sourceActivityTBq = "Source Activity TBq",
  eventNature = "Nature of event"
)

events_sf <- raw_events_df %>% 
  rename(!!!new_names) %>%
  mutate(date = as.Date(date, format = "%d-%b-%y")) %>% 
  drop_na(long, lat) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326)
```

```{r}
events_sf %>% 
  select(date, timeUTC, timeLocal, theftRecovery,
         category, distanceTheftRecovery, geometry) %>% 
  prettify()
```

```{r}
ggplot() +
  geom_sf(data = base_geom_sf, fill = "lightgray") +
  geom_sf(data = events_sf,
          aes(color = theftRecovery),
          show.legend = "point") +
  guides(color = guide_legend(title = NULL)) +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(text = element_text("serif"),
        legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  labs(title = "Raiological Thefts",
       subtitle = str_glue("Mexico, {year1}-{year2}", 
                           year1 = year(min(events_sf$date)),
                           year2 = year(max(events_sf$date))),
       caption = "Courtesy of Francisco \"Paco\" Parada")
```


```{r}
write_rds(events_sf, "data/events_sf.rds")
```


## Municipality Geometries

```{r}
temp_dir <- tempdir()

write(RJSONIO::toJSON(mxmaps::mxmunicipio.topoJSON), 
      file.path(temp_dir, "mxmunicipio.topojson"))

# per https://gist.github.com/diegovalle/993cca843092bf8df692: crs = 4326
municipalities_sf <- read_sf(file.path(temp_dir, "mxmunicipio.topojson"),
                             stringsAsFactors = FALSE,
                             crs = 4326) %>%
  as_tibble() %>%
  rename(INEGI = id) %>% 
  st_as_sf()
```

```{r}
ggplot() +
  geom_sf(data = municipalities_sf,
          aes(fill = INEGI),
          size = 0.01, show.legend = FALSE) +
  guides(color = guide_legend(title = NULL)) +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(text = element_text("serif"),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  labs(title = "Mexican Municipalities",
       caption = "Source: INEGI, courtesy of Diego Valle-Jones")
```

```{r}
write_rds(municipalities_sf, "data/municipalities_sf.rds")
```

# Crime Geometries

```{r}
crimes_sf <- municipalities_sf %>% 
  left_join(crimes_df) %>% 
  drop_na(date)
```

```{r}
crimes_for_plot <- crimes_sf %>% 
  mutate(year = year(date)) %>% 
  st_transform(3857) %>% 
  st_buffer(0) %>% 
  group_by(INEGI) %>% 
  summarise(count = sum(count))

crimes_for_plot %>% 
  ggplot() +
  geom_sf(aes(fill = count),
          color = "transparent") +
  scale_fill_distiller(palette = "RdYlBu", trans = "log10") +
  guides(fill = guide_colorbar(
    title = expression(paste(log[10]~(n[crimes])), parse = TRUE),
    title.position = "top",
    title.hjust = 0.5,
    label.position = "bottom",
    barwidth = grid::unit(10, "cm"))
    ) +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(text = element_text("serif"),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "bottom") +
  labs(title = "Highway Robberies and Vehicle Crimes", 
       subtitle = str_glue("Mexico, {year1}-{year2}", 
                           year1 = year(min(crimes_sf$date)),
                           year2 = year(max(crimes_sf$date))),
       caption = "Source: hoyodecrimen.com, courtesy of Diego Valle-Jones")
```

```{r}
write_rds(crimes_sf, "data/crimes_sf.rds")
```

# Reproducibility

```{r}
sess_info <- devtools::session_info()
```

## Platform

```{r}
sess_info$platform %>% 
  unlist() %>%
  as.matrix() %>% 
  data.frame() %>% 
  rownames_to_column() %>% 
  `colnames<-`(c("setting", "value")) %>% 
  kable()
```

## Packages

```{r}
sess_info$packages %>% 
  kable()
```



